---
title: "Homework 2"
author: Chloe Chah
output: github_document
editor_options: 
  chunk_output_type: console
---



```{r}
library(tidyverse)
library(readxl)
```

## Problem 1 

#### Read the Mr.Trashwheel dataset. 

```{r}

trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
```

#### Read precipitation data! 

```{r}
  precip_2018 = 
   read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
     sheet = "2018 Precipitation", 
     skip = 1
   )%>% 
  janitor::clean_names() %>%
  drop_na (month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
     sheet = "2017 Precipitation", 
     skip = 1
   )%>% 
  janitor::clean_names() %>%
  drop_na (month) %>%
  mutate(year = 2017) %>%
  relocate(year)

```

#### Now combine annual precipitation.

```{r}
  month_df = 
    tibble(
      month = 1:12, 
      month_name = month.name
    )
  
  precip_df = 
    bind_rows(precip_2018, precip_2017)
  
  left_join(precip_df, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data.

## Problem 2 

#### Read in csv file, clean variable names, convert entry variable 

```{r}
   transit_df = read_csv(
     "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
    janitor::clean_names() %>%
    select(line:entry, vending, ada) %>%
    mutate(entry = ifelse(entry == "YES", "True", "False"))
```

#### Description of the NYC Transit dataset 

The dataset contains information related to each entrance and exit for each subway station in New York City. It contains the following variables: `r ls(transit_df)`. The dataset has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. 

In terms of the data cleaning steps, I first read the csv file into R, used the clean_names() function to clean the variable names. Then I used the select function to retain only the desired variables. 

This dataset is not yet tidy because there are redundant columns that seem to describe the same thing. 




## Problem 3 

```{r}
  pols_df = read_csv(
    "./data/pols-month.csv") %>%
    janitor::clean_names() %>%
    separate(mon, sep = "-", into = c("year", "month", "day"))

```

